MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-24 paper.Rmd
Adding FAO and ILO links.

<!-- Define some macros to use later -->
[link-cran]:           https://cran.r-project.org
[link-ieatools]:       https://github.com/MatthewHeun/IEATools
[link-ieaweeb]:        https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
[link-matsindf]:       https://github.com/MatthewHeun/matsindf
[link-mwtools]:        https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfudatabase]:    https://github.com/energyeconomydecoupling/PFUDatabase
[link-pfuaggdatabase]: https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-rclabels]:       https://github.com/MatthewHeun/RCLabels
[link-recca]:          https://github.com/MatthewHeun/Recca
[link-targets]:        https://docs.ropensci.org/targets
[link-tidyverse]:      https://www.tidyverse.org

@@

Several calculation steps are required to create the PFU database. 
Each step is assisted by a new `R` package.
First, the IEA's primary- and final-stage WEEB data must be converted
to the PSUT format, 
a task completed by the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
Second, human and animal muscle work must be calculated from 
[International Labor Organization](https://www.ilo.org/) (ILO) and
[Food and Agriculture Organization](https://www.fao.org/) (FAO) data,
following the methodology of @Steenwyk:2022ww, 
using the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
Third, the IEA's primary- and final-stage ECC data 
are extended to the useful stage by
(a) allocating final stage energy to end-use machines and
(b) multiplying allocated final energy by the 
final-to-useful efficiency of each machine.
This task is accomplished by the
[Recca][link-recca] [@Heun-Recca:2023]
and 
[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) [@Heun-PFUDatabase:2023]
packages.
The pipeline is modular, such that allocation and efficiency data
for a country can be added at any time to improve the database. 
Finally, ECCs must be converted from energy terms to exergy terms. 
This step is assisted by the [Recca][link-recca] package.
The steps to create the PSUT matrices for each country and each year
are accomplished by a
[targets][link-targets] [@Landau:2021aa]
computation pipeline
available in the [PFUDatabase][link-pfudatabase]
package.
A unique feature of the
[PFUDatabase][link-pfudatabase]
pipeline is an exemplar system that allows
analyses to proceed when allocation or efficiency data for a country are unavailable.

@@

| Package | Function |
|:--------|:---------|
| [IEATools][link-idatools]             | Converts IEA data to [matsindf][link-matsindf] format |
| [MWTools][link-mwtools]               | Converts ILO and FAO data to human and animal muscle work information in [matsindf][link-matsindf] format |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | A [targets][link-targets] pipeline to aggregate PSUT matrices |

>>>>

<!-- Define some macros to use later -->
[link-cran]:           https://cran.r-project.org
[link-ieatools]:       https://github.com/MatthewHeun/IEATools
[link-ieaweeb]:        https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:            https://www.ilo.org/
[link-fao]:            https://www.fao.org/
[link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
[link-matsindf]:       https://github.com/MatthewHeun/matsindf
[link-mwtools]:        https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfudatabase]:    https://github.com/energyeconomydecoupling/PFUDatabase
[link-pfuaggdatabase]: https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-rclabels]:       https://github.com/MatthewHeun/RCLabels
[link-recca]:          https://github.com/MatthewHeun/Recca
[link-targets]:        https://docs.ropensci.org/targets
[link-tidyverse]:      https://www.tidyverse.org

@@

Several calculation steps are required to create the PFU database. 
Each step is assisted by a new `R` package.
First, the IEA's primary- and final-stage WEEB data must be converted
to the PSUT format, 
a task completed by the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
Second, human and animal muscle work must be calculated from 
[International Labor Organization][link-ilo] (ILO) and
[Food and Agriculture Organization][link-fao] (FAO) data,
following the methodology of @Steenwyk:2022ww, 
using the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
Third, the IEA's primary- and final-stage ECC data 
are extended to the useful stage by
(a) allocating final stage energy to end-use machines and
(b) multiplying allocated final energy by the 
final-to-useful efficiency of each machine.
This task is accomplished by the
[Recca][link-recca] [@Heun-Recca:2023]
and 
[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) [@Heun-PFUDatabase:2023]
packages.
The pipeline is modular, such that allocation and efficiency data
for a country can be added at any time to improve the database. 
Finally, ECCs must be converted from energy terms to exergy terms. 
This step is assisted by the [Recca][link-recca] package.
The steps to create the PSUT matrices for each country and each year
are accomplished by a
[targets][link-targets] [@Landau:2021aa]
computation pipeline
available in the [PFUDatabase][link-pfudatabase]
package.
A unique feature of the
[PFUDatabase][link-pfudatabase]
pipeline is an exemplar system that allows
analyses to proceed when allocation or efficiency data for a country are unavailable.

@@

| Package | Function |
|:--------|:---------|
| [IEATools][link-ieatools]             | Converts IEA data to [matsindf][link-matsindf] format |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to human and animal muscle work in [matsindf][link-matsindf] format |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | A [targets][link-targets] pipeline to aggregate PSUT matrices |
